---
title: "Lab 4"
author: ""
date: ""
output: 
  html_document:
    toc: true
    number_sections: false
    toc_float: true
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

## Packages

```{r}
library(tidyverse)
```

## NC absentee voting data

>Voter information includes name, demographics, mailing address, precinct, and 
district data. Ballot information for voting by mail includes request date, 
send date, return date, return status (e.g., ACCEPTED, SPOILED, etc.), and 
mailing address(es). Ballot information for one-stop early voting includes 
voting date (called return date), early voting site, and same-day registration 
information. For full list of variables included, see 
[file layout](https://s3.amazonaws.com/dl.ncsbe.gov/ENRS/layout_absentee.txt).
<br/><br/>
*Source*: https://www.ncsbe.gov/results-data/absentee-data

```{r get_data}
base_url <- "https://s3.amazonaws.com/"
zip_url <-str_c(base_url, "dl.ncsbe.gov/ENRS/2020_11_03/absentee_20201103.zip")
download.file(url = zip_url, destfile = "absentee_votes.zip")
unzip("absentee_votes.zip")
absentee <- read_csv("absentee_20201103.csv")
```

## Visualizations

As you work on creating the plots, feel free to be creative with color choices
and other aesthetics.

Create a bar plot of the number of ballot received by ballot send date.

See `images/visualizaiton_1.png`.

```{r visualizaiton_1}
absentee %>%
  group_by(ballot_send_dt) %>%
  count %>%
  arrange(n) %>%
  ggplot(aes(ballot_send_dt, n)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Absentee Votes by Ballot Send Date", x = "Ballot Send Date", y = "Absentee Ballots")
```

Create a bar plot showing the number of ballots received on 09/04/2020 by 
county. Only show the top 20 counties as determined by number of
ballots received on 09/04/2020.

See `images/visualizaiton_2.png`.

```{r visualizaiton_2}
absentee %>%
  filter(ballot_send_dt == "09/04/2020") %>%
  count(county_desc) %>%
  arrange(desc(n)) %>%
  head(20) %>%
  ggplot(aes(reorder(county_desc, n), n)) + 
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Absentee Votes on 9/4/20 by County", x = "County", y = "Absentee Ballots")
```

Create a histogram based on the proportion of accepted ballots on a per
county basis.

See `images/visualizaiton_3.png`.

```{r visualizaiton_3}
absentee %>%
  group_by(county_desc) %>%
  summarize(n = n(), prop = length(ballot_rtn_status[ballot_rtn_status == "ACCEPTED"])/n)
  arrange(desc(prop)) %>%
  ggplot(aes(prop)) +
  geom_histogram()
```
  
```{r}
absentee %>%
  group_by(county_desc) %>%
  count(ballot_rtn_status) %>%
  mutate(prop = n/sum(n)) %>%
  ungroup() %>%
  filter(ballot_rtn_status == "ACCEPTED") %>%
  ggplot(aes(x = prop)) +
  geom_histogram(binwidth = 0.02)

absentee %>%
  group_by(county_desc) %>%
  count(ballot_rtn_status) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup() %>%
  filter(ballot_rtn_status == "ACCEPTED") %>%
  ggplot(aes(x = proportion)) +
  geom_histogram(binwidth = 0.02) +
  labs(x = "Proportion of ballots accepted", y = "Count") +
  theme_minimal(base_size = 16)

```

Create comparison box plots based on the age of absentee/mail-in voters
from the parties "DEM", "REP", "UNA" broken down by race.

See `images/visualizaiton_4.png`.

```{r visualizaiton_4}
ggplot(absentee, aes(x = age)) +
  geom_boxplot(aes(color = race)) +
  facet_wrap(~ voter_party_code)
```
Extra visualizations: 

Largest Democratic counties in NC: 

```{r}
absentee %>%
  filter(voter_party_code == "DEM") %>%
  group_by(county_desc) %>%
  count(voter_party_code) %>%
  arrange(desc(n)) %>%
  head(20) %>%
  ggplot(aes(reorder(county_desc, n), n)) +
  geom_bar(stat = "identity") +
  coord_flip()
```
Top democratic counties proportion of accepted ballots: 

Almost 100% of absentee ballots have been accepted for top Democratic counties. However, New Hanover has one of the highest rates of ballot acceptance at 83%. 

```{r}
absentee %>%
  filter(voter_party_code == "DEM") %>%
  group_by(county_desc) %>%
  count(voter_party_code) %>% 
  arrange(desc(n)) 
```

```{r}
absentee %>%
  filter(county_desc %in% c("MECKLENBURG", "WAKE", "DURHAM", "CUMBERLAND", "ORANGE", "HENDERSON", "PITT", "ALAMANCE", "NEW HANOVER", "BRUNSWICK")) %>%
  group_by(county_desc) %>%
  count(ballot_rtn_status) %>%
  mutate(prop = n/sum(n), ballot_rtn_status) %>%
  ungroup() %>%
  filter(ballot_rtn_status == "ACCEPTED")
```

Which counties have the highest rates of ballots not being accepted? 

New Hanover has a very high rate of not being accepted. 

```{r}
absentee %>%
  group_by(county_desc) %>%
  count(ballot_rtn_status) %>%
  mutate(prop = n/sum(n)) %>%
  ungroup() %>%
  filter(ballot_rtn_status != "ACCEPTED") %>%
  arrange(desc(prop))
```

